Enhancing Underwater Imagery using Generative Adversarial Networks

### 2017-09-16 files/paper/tex/abstract.tex
paper complete, first full draft.

\begin{abstract}
Autonomous underwater vehicles (AUVs) rely on a variety of sensors -- acoustic, inertial and visual -- for intelligent decision 
making. Due
to its non-intrusive, passive nature, and high information content, vision is an attractive sensing modality, particularly at 
shallower
depths. However, factors such as light refraction and absorption, suspended particles in the water, and color distortion affect 
the quality
of visual data, resulting in noisy and distorted images. AUVs that rely on visual sensing thus face difficult challenges, and 
consequently
exhibit poor performance on vision-driven tasks. This paper proposes a method to improve the quality of visual underwater scenes 
using
Generative Adversarial Networks (GANs), with the goal of improving input to vision-driven behaviors further down the autonomy
pipeline. Furthermore, we show how recently proposed methods are able to generate a dataset for the purpose of such underwater 
image
reconstruction\starnote{Reconstruction meaning what we are doing here, right? So this is a contribution too, to show how CycleGAN 
is used to
generate the dataset, correct?}. For any visually-guided underwater robots, this improvement can result in increased safety and 
reliability
through robust visual perception. To that effect, we present quantitative and qualitative data which demonstrates that images 
corrected
through the proposed approach generate more visually appealing images, and also provide increased accuracy for an underwater diver 
tracking
algorithm.
\end{abstract}

>>>>

\begin{abstract}
Autonomous underwater vehicles (AUVs) rely on a variety of sensors -- acoustic, inertial and visual -- for intelligent decision 
making. Due to its non-intrusive, passive nature, and high information content, vision is an attractive sensing modality, 
particularly at shallower depths. However, factors such as light refraction and absorption, suspended particles in the water, and 
color distortion affect the quality of visual data, resulting in noisy and distorted images. AUVs that rely on visual sensing thus 
face difficult challenges, and consequently exhibit poor performance on vision-driven tasks. This paper proposes a method to 
improve the quality of visual underwater scenes using Generative Adversarial Networks (GANs), with the goal of improving input to 
vision-driven behaviors further down the autonomy pipeline. Furthermore, we show how recently proposed methods are able to 
generate a dataset for the purpose of such underwater image restoration\nostarnote{Reconstruction meaning what we are doing here, 
right? So this is a contribution too, to show how CycleGAN is used to generate the dataset, correct?}. For any visually-guided 
underwater robots, this improvement can result in increased safety and reliability through robust visual perception. To that 
effect, we present quantitative and qualitative data which demonstrates that images corrected through the proposed approach 
generate more visually appealing images, and also provide increased accuracy for an underwater diver tracking algorithm.
\end{abstract}

### 2017-09-16 files/paper/tex/conclusions.tex
Minor edits.
paper complete, first full draft.

\section{Conclusion}

Future work: Getting a better and larger dataset. I think that would help a great deal. Also data augmentation
after CycleGAN, such as gaussian blurring, noise, particle effect, more lighting effects etc.

\section*{Acknowledgment}
Anyone we wish to acknowledge?

>>>>

\section{Conclusion}
This paper presents an approach for enhancing underwater color images through the use of generative adversarial networks. We 
demonstrate the use of CycleGAN to generate dataset of paired images to provide a training set for the proposed restoration 
model. Quantitative and qualitative results demonstrate the effectiveness of this method, and using a diver tracking algorithm on 
corrected images of scuba divers show higher accuracy compared to the uncorrected image sequence.

Future work will focus on creating a larger and more diverse dataset from underwater objects, thus making the network more 
generalizable. Augmenting the data generated by CycleGAN with noise such as particle and lighting effects would improve 
the diversity of the dataset. We also intend to investigate a number of different quantitative performance metrics to evaluate 
our method.

\section*{Acknowledgment}
The authors are grateful to Oliver Hennigh for his implementation of the Gradient Difference Loss measure.

### 2017-09-16 files/paper/tex/experiments.tex
paper complete, first full draft.

\subsection{Datasets}
We used several subsets of Imagenet~\cite{deng2009imagenet} for training and evaluation of our methods. We also evaluate a 
tracking
algorithm on a video of scuba divers taken from Youtube \footnote{https://www.youtube.com/watch?v=QmRFmhILd5o}. Subsets of 
Imagenet
containing underwater images were selected for the training of CycleGAN, and manually separated into two classes based on visual
inspection. We let $X$ be the set of underwater images with no distortion, and $Y$ be the set of underwater images with 
distortion. $X$
contained 6143 images, and $Y$ contained 1817 images. We then trained CycleGAN to learn the mapping $F: X \rightarrow Y$, such 
that images
from $X$ appeared to have come from $Y$. Finally, our image pairs for training data were generated by distorting all images in $X$ 
with
$F$. Figure~\ref{fig:cgan_samples} shows sample training pairs. When comparing with CycleGAN, we used a test set of 56 images 
acquired from
Flickr\texttrademark .

@@

\begin{figure}
\centering
\begin{tabular}{p{1.7cm} p{1.7cm} p{1.7cm}}

@@

\end{tabular}
\caption{Samples from our Imagenet testing set. The network can both recover color and also correct color if a small amount is 
present.}
\label{fig:test_samples}
\end{figure}

@@

\subsection{Training and Inference Performance}
In all of our experiments, we use $\lambda_{1} = 100$, $\lambda_{GP} = 10$, batch size of 32, and the Adam Optimizer 
\cite{kingma2014adam} with learning rate $1e-4$. Following WGAN-GP, the discriminator is updated $n$ times for every update of the 
generator, where $n = 5$. For UGAN-P, we set $\lambda_{2} = 1.0$ and $\alpha = 1$. Our implementation was done using the 
Tensorflow library \cite{abadi2016tensorflow}. \footnote{Code is available at 
\url{https://github.com/cameronfabbri/Underwater-Color-Correction}} All networks were trained from scratch on a GTX 1080 for 100 
epochs. Inference on the GPU takes on average $ 0.0138s$, which is about 72 Frames Per Second (FPS). On a CPU (Intel Core 
i7-5930K), inference takes on average $ 0.1244s$, which is about 8 FPS. We find both of these measures acceptable for underwater 
tasks.

\subsection{Comparison to CycleGAN}
It is important to note that during the process of learning a mapping $F: X \rightarrow Y$, CycleGAN also learns a
mapping $G: Y \rightarrow X$. Here we give a comparison to our methods. We use the Canny edge
detector \cite{canny1986computational} to provide a color agnostic evaluation of the images, as the original contain
distorted colors and cannot be compared back to as ground truth. Due to the fact that restoring color information
should not alter the overall structure of the image, we measure the distance in the image space between the edges
found in the original and generated images. Figure ~\ref{fig:canny_samples} shows the original images and results from
edge detection. Table 1 provides the measurements from Figure ~\ref{fig:canny_samples}, as well as the average over our
entire Flickr\texttrademark ~dataset. Both UGAN and UGAN-P are consistently closer in the image space to the original
than that of CycleGAN, suggesting noise due to blur. Next, we evaluate this noise explicitly.

We explore the artifacts of content loss, as seen in Figure ~\ref{fig:zoom}. In particular, we compare local statistics
of the highlighted image patches, where each image patch is resized to $64 \times 64$. We use the GDL
\cite{mathieu2015deep} from (\ref{gdl_eq}) as a sharpness measure. A lower GDL measure implies a smoother transition
between pixels, as a noisy image would have large jumps in the image's gradient, leading to a higher score. \starnote{I THINK. I would double check, bottom of page 4: https://arxiv.org/pdf/1511.05440.pdf}.
As seen in Table \ref{fig:gdl_tbl}, the GDL is lower for both UGAN and UGAN-P. Interestingly, UGAN consistently has a
lower score than UGAN-P, despite UGAN-P explicitly accounting for this metric in the objective function. Reasoning for
this is left for our future work.

Another metric we can use to compare image patches are the mean and standard deviation of a patch. The standard
deviation gives us a sense of blurriness because it defines how far the data deviates from the mean. A lower standard
deviation means that the data is closer to the mean \starnote{not sure if we exactly need to say that because I assume the reviewers will know what it is}. In the case of images, this would suggest a blurring effect due to the data being
more clustered toward one pixel value. Table \ref{fig:mean_tbl} shows the mean and standard deviations for the local
image patches seen in Figure \ref{fig:zoom}. Despite qualitative evaluation showing our methods are much sharper,
quantitatively they show only slight improvement. Other metrics such as entropy are left as future work.

% I would like this table to match up with the figure showing edge detections. I wasn't able to
% get letters on the left side at each row though.
\begin{table}
\centering
\caption{Distances in image space}
\begin{tabular}{| c | c | c | c |}
   \hline
   Row/Method & CycleGAN & \textbf{UGAN} & \textbf{UGAN-P} \\ \hline
   A          & 116.45 & 85.71  & 86.15  \\ \hline
   B          & 114.49 & 97.92  & 101.01 \\ \hline
   C          & 120.84 & 96.53  & 97.57  \\ \hline
   D          & 129.27 & 108.90 & 110.50 \\ \hline
   Mean       & 111.60 & 94.91  & 96.51 \\ \hline
\end{tabular}
\end{table}

@@

\begin{figure}
\centering
\begin{tabular}{p{1.7cm} p{1.7cm} p{1.7cm} p{1.7cm}}

@@

\end{tabular}
\caption{Running the Canny Edge Detector on sample images. Both variants of UGAN contain less noise than CycleGAN,
and are closer in the image space to the original. For each pair, the top row is the input image, and bottom row
the result of the edge detector.}
\label{fig:canny_samples}
\end{figure}

@@

\begin{table}[ht]
\footnotesize
\centering
\caption{Gradient Difference Loss Metrics}
\begin{tabular}{| c | c | c | c | }
   \hline
   Patch/Method & CycleGAN & \textbf{UGAN} & \textbf{UGAN-P} \\ \hline
   Red    & 11.53 & 9.39 & 10.93  \\ \hline
   Blue   & 7.52  & 4.83 &  5.50\\ \hline
   Green  & 4.15  & 3.18 & 3.25 \\ \hline
   Orange & 6.72  & 5.65 & 5.79 \\ \hline
\end{tabular}
\label{fig:gdl_tbl}
\end{table}

\begin{table}[ht]
\footnotesize
\centering
\caption{Mean and Standard Deviation Metrics}
\begin{tabular}{| c | c | c | c | c | }
   \hline
   Patch/Method & Original & CycleGAN & \textbf{UGAN} & \textbf{UGAN-P} \\ \hline
   Red & 0.43 $\pm$ 0.23 & 0.42 $\pm$ 0.22 & 0.44 $\pm$ 0.23 & 0.45 $\pm$ 0.25 \\ \hline
   Blue & 0.51 $\pm$ 0.18 & 0.57 $\pm$ 0.17 & 0.57 $\pm$ 0.17 & 0.57 $\pm$ 0.17 \\ \hline
   Green & 0.36 $\pm$ 0.17 & 0.36 $\pm$ 0.14 & 0.37 $\pm$ 0.17 & 0.36 $\pm$ 0.17 \\ \hline
   Orange & 0.3 $\pm$ 0.15 & 0.25 $\pm$ 0.12 & 0.26 $\pm$ 0.13 & 0.27 $\pm$ 0.14 \\ \hline
\end{tabular}
\label{fig:mean_tbl}
\end{table}

@@

%Flippers of a human diver typically oscillate at frequencies between $1$ and $2$ Hz, which produces periodic intensity variations 
in the image-space over time. These variations correspond to distinct signatures in the frequency-domain (high-amplitude spectra 
at $1$-$2$Hz), which can be used for reliable detection.

>>>>

\subsection{Datasets}
We used several subsets of Imagenet~\cite{deng2009imagenet} for training and evaluation of our methods. We also evaluate a 
frequency- and spatial-domain diver-tracking algorithm on a video of scuba divers taken from YouTube\texttrademark\ 
\footnote{https://www.youtube.com/watch?v=QmRFmhILd5o}. Subsets of 
Imagenet
containing underwater images were selected for the training of CycleGAN, and manually separated into two classes based on visual
inspection. We let $X$ be the set of underwater images with no distortion, and $Y$ be the set of underwater images with 
distortion. $X$
contained 6143 images, and $Y$ contained 1817 images. We then trained CycleGAN to learn the mapping $F: X \rightarrow Y$, such 
that images
from $X$ appeared to have come from $Y$. Finally, our image pairs for training data were generated by distorting all images in $X$ 
with
$F$. Figure~\ref{fig:cgan_samples} shows sample training pairs. When comparing with CycleGAN, we used a test set of 56 images 
acquired from
Flickr\texttrademark .

@@

\begin{figure}[th]
\centering
\begin{tabular}{p{1.7cm} p{1.7cm} p{1.75cm}}

@@

\end{tabular}
\caption{Samples from our ImageNet testing set. The network can both recover color and also correct color if a small amount is 
present.}
\label{fig:test_samples}
\end{figure}

@@

\subsection{Comparison to CycleGAN}
It is important to note that during the process of learning a mapping $F: X \rightarrow Y$, CycleGAN also learns a mapping $G: Y 
\rightarrow X$. Here we give a comparison to our methods. We use the Canny edge detector \cite{canny1986computational} to provide 
a color agnostic evaluation of the images, as the original contain distorted colors and cannot be compared back to as ground 
truth. Due to the fact that restoring color information should not alter the overall structure of the image, we measure the 
distance in the image space between the edges found in the original and generated images. Figure ~\ref{fig:canny_samples} shows 
the original images and results from edge detection. Table~\ref{tab:one} provides the measurements from Figure 
~\ref{fig:canny_samples}, as well as the average over our entire Flickr\texttrademark\ dataset. Both UGAN and UGAN-P are 
consistently closer in the image space to the original than that of CycleGAN, suggesting noise due to blur. Next, we evaluate this 
noise explicitly.

We explore the artifacts of content loss, as seen in Figure ~\ref{fig:zoom}. In particular, we compare local statistics of the 
highlighted image patches, where each image patch is resized to $64 \times 64$. We use the GDL \cite{mathieu2015deep} from 
(\ref{gdl_eq}) as a sharpness measure. A lower GDL measure implies a smoother transition between pixels, as a noisy image would 
have large jumps in the image's gradient, leading to a higher score.\nostarnote{I THINK. I would double check, bottom of page 4: 
https://arxiv.org/pdf/1511.05440.pdf.} As seen in Table \ref{fig:gdl_tbl}, the GDL is lower for both UGAN and UGAN-P. 
Interestingly, UGAN consistently has a lower score than UGAN-P, despite UGAN-P explicitly accounting for this metric in the 
objective function. Reasoning for this is left for our future work.

Another metric we use to compare image patches are the mean and standard deviation of a patch. The standard deviation gives us a 
sense of blurriness because it defines how far the data deviates from the mean.\nostarnote{ A lower standard deviation means that 
the data is closer to the mean . not sure if we exactly need to say that because I assume the reviewers will know what it is.} In 
the case of images, this would suggest a blurring effect due to the data being more clustered toward one pixel value. Table 
\ref{fig:mean_tbl} shows the mean and standard deviations of the RGB values for the local image patches seen in Figure 
\ref{fig:zoom}. Despite qualitative evaluation showing our methods are much sharper, quantitatively they show only slight 
improvement. Other metrics such as entropy are left as future work.

% I would like this table to match up with the figure showing edge detections. I wasn't able to
% get letters on the left side at each row though.
\begin{table}
\centering
\caption{Distances in image space}
\begin{tabular}{| c | c | c | c |}
   \hline
   Row/Method & CycleGAN & \textbf{UGAN} & \textbf{UGAN-P} \\ \hline
   A          & 116.45 & 85.71  & 86.15  \\ \hline
   B          & 114.49 & 97.92  & 101.01 \\ \hline
   C          & 120.84 & 96.53  & 97.57  \\ \hline
   D          & 129.27 & 108.90 & 110.50 \\ \hline
   Mean       & 111.60 & 94.91  & 96.51 \\ \hline
\end{tabular}
\label{tab:one}
\end{table}

@@

% \begin{figure*}[h]
% \centering
% \begin{tabular}{p{.4cm} p{1.7cm} p{1.7cm} p{1.7cm} p{1.7cm}}
%   
%    Set & ~\quad Original & ~CycleGAN & ~\quad \textbf{UGAN} & \quad \textbf{UGAN-P} \\
% 
%    \multirow{2}{*}{A} & \includegraphics[width=0.8in]{1_original} &
%    \includegraphics[width=0.8in]{1_cimg} &
%    \includegraphics[width=0.8in]{1_u0img} &
%    \includegraphics[width=0.8in]{1_u1img} \\ [-1ex]
%    \includegraphics[width=0.8in]{1_oedges} &
%    \includegraphics[width=0.8in]{1_cedges} &
%    \includegraphics[width=0.8in]{1_u0edges} &
%    \includegraphics[width=0.8in]{1_u1edges} \\
% 
%    \multirow{2}{*}{B} & \includegraphics[width=0.8in]{2_original} &
%    \includegraphics[width=0.8in]{2_cimg} &
%    \includegraphics[width=0.8in]{2_u0img} &
%    \includegraphics[width=0.8in]{2_u1img} \\ [-1ex]
%    \includegraphics[width=0.8in]{2_oedges} &
%    \includegraphics[width=0.8in]{2_cedges} &
%    \includegraphics[width=0.8in]{2_u0edges} &
%    \includegraphics[width=0.8in]{2_u1edges} \\
% 
%    \multirow{2}{*}{C} & \includegraphics[width=0.8in]{3_original} &
%    \includegraphics[width=0.8in]{3_cimg} &
%    \includegraphics[width=0.8in]{3_u0img} &
%    \includegraphics[width=0.8in]{3_u1img} \\ [-1ex]
%    \includegraphics[width=0.8in]{3_oedges} &
%    \includegraphics[width=0.8in]{3_cedges} &
%    \includegraphics[width=0.8in]{3_u0edges} &
%    \includegraphics[width=0.8in]{3_u1edges} \\
%    
%    \multirow{2}{*}{D} & \includegraphics[width=0.8in]{4_original} &
%    \includegraphics[width=0.8in]{4_cimg} &
%    \includegraphics[width=0.8in]{4_u0img} &
%    \includegraphics[width=0.8in]{4_u1img} \\ [-1ex]
%    \includegraphics[width=0.8in]{4_oedges} &
%    \includegraphics[width=0.8in]{4_cedges} &
%    \includegraphics[width=0.8in]{4_u0edges} &
%    \includegraphics[width=0.8in]{4_u1edges} \\
% 
% \end{tabular}
% \caption{Running the Canny Edge Detector on sample images. Both variants of UGAN contain less noise than CycleGAN,
% and are closer in the image space to the original. For each pair, the top row is the input image, and bottom row
% the result of the edge detector.}
% \label{fig:canny_samples}
% \end{figure*}

\begin{figure}[h]
\centering
\begin{tabular}{p{1.7cm} p{1.7cm} p{1.7cm} p{1.75cm}}

@@

\end{tabular}
\caption{Running the Canny Edge Detector on sample images. Both variants of UGAN contain less noise than CycleGAN,
and are closer in the image space to the original. For each pair, the top row is the input image, and bottom row
the result of the edge detector. The figure depicts four different sets of images, successively labeled A to D from top to 
bottom. See Table~\ref{tab:one}.}
\label{fig:canny_samples}
\vspace{-5mm}
\end{figure}

@@

\begin{table}[ht]
\footnotesize
\centering
\caption{Gradient Difference Loss Metrics}
\begin{tabular}{| c | c | c | c | }
   \hline
   \multicolumn{1}{|p{.7cm}|}{\centering Method/ \\ Patch} & CycleGAN & \textbf{UGAN} & \textbf{UGAN-P} \\ \hline
   Red    & 11.53 & 9.39 & 10.93  \\ \hline
   Blue   & 7.52  & 4.83 &  5.50\\ \hline
   Green  & 4.15  & 3.18 & 3.25 \\ \hline
   Orange & 6.72  & 5.65 & 5.79 \\ \hline
\end{tabular}
\label{fig:gdl_tbl}
\end{table}

\begin{table}[ht]
\footnotesize
\centering
\caption{Mean and Standard Deviation Metrics}
\begin{tabular}{| c | c | c | c | c | }
   \hline
   \multicolumn{1}{|p{.7cm}|}{\centering Method/ \\ Patch} & Original & CycleGAN & \textbf{UGAN} & \textbf{UGAN-P} \\ 
\hline
   Red & 0.43 $\pm$ 0.23 & 0.42 $\pm$ 0.22 & 0.44 $\pm$ 0.23 & 0.45 $\pm$ 0.25 \\ \hline
   Blue & 0.51 $\pm$ 0.18 & 0.57 $\pm$ 0.17 & 0.57 $\pm$ 0.17 & 0.57 $\pm$ 0.17 \\ \hline
   Green & 0.36 $\pm$ 0.17 & 0.36 $\pm$ 0.14 & 0.37 $\pm$ 0.17 & 0.36 $\pm$ 0.17 \\ \hline
   Orange & 0.3 $\pm$ 0.15 & 0.25 $\pm$ 0.12 & 0.26 $\pm$ 0.13 & 0.27 $\pm$ 0.14 \\ \hline
\end{tabular}
\label{fig:mean_tbl}
\end{table}

@@

\subsection{Training and Inference Performance}
In all of our experiments, we use $\lambda_{1} = 100$, $\lambda_{GP} = 10$, batch size of 32, and the Adam Optimizer 
\cite{kingma2014adam} with learning rate $1e-4$. Following WGAN-GP, the discriminator is updated $n$ times for every update of the 
generator, where $n = 5$. For UGAN-P, we set $\lambda_{2} = 1.0$ and $\alpha = 1$. Our implementation was done using the 
Tensorflow library \cite{abadi2016tensorflow}. \footnote{Code is available at 
\url{https://github.com/cameronfabbri/Underwater-Color-Correction}} All networks were trained from scratch on a GTX 1080 for 100 
epochs. Inference on the GPU takes on average $ 0.0138s$, which is about 72 Frames Per Second (FPS). On a CPU (Intel Core 
i7-5930K), inference takes on average $ 0.1244s$, which is about 8 FPS. In both cases, the input images have dimensions 
$256\times 256\times 3$. We find both of these measures acceptable for underwater tasks.

%Flippers of a human diver typically oscillate at frequencies between $1$ and $2$ Hz, which produces periodic intensity variations 
% in the image-space over time. These variations correspond to distinct signatures in the frequency-domain (high-amplitude spectra 
% at $1$-$2$Hz), which can be used for reliable detection.

### 2017-09-16 files/paper/tex/introduction.tex
Minor edits.
paper complete, first full draft.

\section{Introduction}
%TODO - talk about having to go back to the same location if you want to get good/bad pairs of the ``same'' image.\starnote{I 
assume Cam will
%do this?}
Underwater robotics has been a steadily growing subfield of autonomous field robotics, assisted by the advent of novel platforms,
sensors and propulsion mechanisms. While autonomous underwater vehicles are often equipped with a variety of sensors, visual 
sensing is an
attractive option because of its non-intrusive, passive, and energy efficient nature. The monitoring of coral reefs 
\cite{shkurti2012multi},
deep ocean exploration \cite{whitcomb2000advances}, and mapping of the seabed~\cite{bingham2010robotic} are a number of tasks 
where
visually-guided AUVs and ROVs (Remotely Operated Vehicles) have seen widespread use. Use of these robots ensures humans are not 
exposed to
the hazards of underwater exploration, as they no longer need to venture to the depths (which was how such tasks were carried out 
in the
past). Despite the advantages of using vision, underwater environments pose unique challenges to visual sensing, as phenomena such 
as light
refraction, absorption and scattering from suspended particles can greatly affect the optics of light. As an example, because red
wavelengths are quickly absorbed by water, images tend to have a green or blue hue to them. As one goes deeper, this effect 
worsens, as more
and more red hue is absorbed. This distortion is extremely non-linear in nature, and is affected by a large number of factors, 
such as the
amount of light present (overcast versus sunny, operational depth), amount of particles in the water, time of day, and the camera 
being
used. This may cause difficulty in tasks such as segmentation, tracking, or classification due to their indirect or direct use of 
color.

@@

\begin{figure}
\centering
\begin{tabular}{p{1.7cm} p{1.7cm} p{1.7cm} p{1.7cm}}

   \includegraphics[width=0.8in]{n01496331_7428_f1} &
   \includegraphics[width=0.8in]{n01496331_16340_f1} &
   \includegraphics[width=0.8in]{n01914609_5148_f1} &
   \includegraphics[width=0.8in]{robot_f1} \\
   \includegraphics[width=0.8in]{n01664065_30279_f1} &
   \includegraphics[width=0.8in]{n01917289_5711_f1} &
   \includegraphics[width=0.8in]{n01917289_4087_f1} &
   \includegraphics[width=0.8in]{n02607072_10395_f1} \\

\end{tabular}
\label{fig:samples}
\caption{Sample underwater images with natural and man-made artifacts (which in this case is our underwater robot) displaying the 
diversity of distortion that can occur. With the varying camera-to-object distances in the images, the distortion and loss of 
color varies between the different images.}
\end{figure}

Deep neural networks have been shown to be powerful non-linear function approximators, especially in the field of vision 
\cite{krizhevsky2012imagenet}. Often times, these networks require large amounts of data, either labeled or paired with
ground truth. For the problem of automatically colorizing grayscale images \cite{zhang2016colorful}, paired training
data is
readily available due to the fact that any color image can be converted to black and white. However, underwater images distorted 
by either
color or some other phenomenon lack ground truth, which is a major hindrance towards adopting a similar approach for correction. 
This paper proposes a technique based on Generative Adversarial Networks (GANs) to improve the quality of visual underwater scenes 
with the goal of improving the performance of vision-driven behaviors for autonomous underwater robots.
%\starnote{Summarize the approach here. Then talk about datasets as you have done already.}
We use the recently proposed CycleGAN~\cite{zhu2017unpaired} approach, which learns to translate an image from any arbitrary 
domain $X$ to another arbitrary domain $Y$ \textit{without} image pairs, as a way to generate a paired dataset.
By letting $X$ be a set of undistorted underwater images\starnote{need to say where we find the undistorted ones}, and
$Y$ be a set of distorted underwater images, we can generate an image that appears to be underwater while retaining
ground truth.

>>>>

\section{Introduction}
%TODO - talk about having to go back to the same location if you want to get good/bad pairs of the ``same'' image.\starnote{I 
%assume Cam will %do this?}
Underwater robotics has been a steadily growing subfield of autonomous field robotics, assisted by the advent of novel platforms,
sensors and propulsion mechanisms. While autonomous underwater vehicles are often equipped with a variety of sensors, visual 
sensing is an
attractive option because of its non-intrusive, passive, and energy efficient nature. The monitoring of coral reefs 
\cite{shkurti2012multi},
deep ocean exploration \cite{whitcomb2000advances}, and mapping of the seabed~\cite{bingham2010robotic} are a number of tasks 
where
visually-guided AUVs and ROVs (Remotely Operated Vehicles) have seen widespread use. Use of these robots ensures humans are not 
exposed to
the hazards of underwater exploration, as they no longer need to venture to the depths (which was how such tasks were carried out 
in the
past). Despite the advantages of using vision, underwater environments pose unique challenges to visual sensing, as phenomena such 
as light
refraction, absorption and scattering from suspended particles can greatly affect the optics of light. As an example, because red
wavelengths are quickly absorbed by water, images tend to have a green or blue hue to them. As one goes deeper, this effect 
worsens, as more
and more red hue is absorbed. This distortion is extremely non-linear in nature, and is affected by a large number of factors, 
such as the
amount of light present (overcast versus sunny, operational depth), amount of particles in the water, time of day, and the camera 
being
used. This may cause difficulty in tasks such as segmentation, tracking, or classification due to their indirect or direct use of 
color.

@@

\begin{figure}
\vspace{2mm}
\centering
\begin{tabular}{p{1.7cm} p{1.7cm} p{1.7cm} p{1.7cm}}

   \includegraphics[width=0.8in]{n01496331_7428_f1} &
   \includegraphics[width=0.8in]{n01496331_16340_f1} &
   \includegraphics[width=0.8in]{n01914609_5148_f1} &
   \includegraphics[width=0.8in]{robot_f1} \\
   \includegraphics[width=0.8in]{n01664065_30279_f1} &
   \includegraphics[width=0.8in]{n01917289_5711_f1} &
   \includegraphics[width=0.8in]{n01917289_4087_f1} &
   \includegraphics[width=0.8in]{n02607072_10395_f1} \\
\end{tabular}
\label{fig:samples}
\caption{Sample underwater images with natural and man-made artifacts (which in this case is our underwater robot) displaying the 
diversity of distortion that can occur. With the varying camera-to-object distances in the images, the distortion and loss of 
color varies between the different images.}
\vspace{-5mm}
\end{figure}

Deep neural networks have been shown to be powerful non-linear function approximators, especially in the field of vision 
\cite{krizhevsky2012imagenet}. Often times, these networks require large amounts of data, either labeled or paired with
ground truth. For the problem of automatically colorizing grayscale images \cite{zhang2016colorful}, paired training
data is
readily available due to the fact that any color image can be converted to black and white. However, underwater images distorted 
by either
color or some other phenomenon lack ground truth, which is a major hindrance towards adopting a similar approach for correction. 
This paper proposes a technique based on Generative Adversarial Networks (GANs) to improve the quality of visual underwater scenes 
with the goal of improving the performance of vision-driven behaviors for autonomous underwater robots.
%\starnote{Summarize the approach here. Then talk about datasets as you have done already.}
We use the recently proposed CycleGAN~\cite{zhu2017unpaired} approach, which learns to translate an image from any arbitrary 
domain $X$ to another arbitrary domain $Y$ \textit{without} image pairs, as a way to generate a paired dataset.
By letting $X$ be a set of undistorted underwater images\nostarnote{need to say where we find the undistorted ones}, and
$Y$ be a set of distorted underwater images, we can generate an image that appears to be underwater while retaining
ground truth.

### 2017-09-16 files/paper/tex/methodology.tex
Editing stuff for final version.
paper complete, first full draft.

\subsection{Adversarial Networks}
\nostarnote{check this paragraph as I added some more text}In machine learning literature, Generative Adversarial Networks 
(GANs)~\cite{goodfellow2014generative} represent a class of generative models based on game theory in which a generator network 
competes against an adversary. From a classification perspective, the generator network produces instances which actively attempt 
to `fool' the discriminator network. The goal is for the discriminator network to be able to distinguish between `true' 
instances coming from the dataset and `false' instances produced by the generator network.
%, and thus improve its classification accuracy.
In our case, conditioned on an image $I^D$, the generator is trained to produce an image to try and fool the 
discriminator, which is trained to distinguish between distorted and non-distorted underwater images. In the original
GAN formulation, our goal is to solve the minimax problem:

\begin{figure}
\centering
\begin{tabular}{p{1.7cm} p{1.7cm} p{1.7cm} p{1.7cm}}

@@

\end{tabular}
\caption{Paired samples of ground truth and distorted images generated by CycleGAN. Top row: Ground truth.
Bottom row: Generated samples.}
\label{fig:cgan_samples}
\end{figure}

@@

Note for simplicity in notation, we will further omit $I^C \sim p_{train}$ and $I^D \sim p_{gen}$. In this formulation, 
the discriminator is hypothesized as a classifier with a sigmoid cross-entropy loss function, which in practice may lead
to issues  such as the vanish gradient and mode collapse. As shown by \cite{arjovsky2017towards}, as
the discriminator improves, the gradient of the generator vanishes, making it difficult or impossible to train. Mode
collapse occurs when the generator ``collapses'' onto a single point, fooling the discriminator with only one instance
\starnote{Such as when you are training a GAN to generate digits from MNIST, but it only generates a 7, when in reality
you'd want it to generate a diverse amount of all the digits. Not sure if we should expand more.}.
There have been a number of recent methods which hypothesize a different loss function for the discriminator
\cite{mao2016least,arjovsky2017wasserstein,gulrajani2017improved,zhao2016energy}. We focus on the Wasserstein GAN
(WGAN) \cite{arjovsky2017wasserstein} formulation, which proposes to use the Earth-Mover or \textit{Wasserstein-1}
distance $W$ by constructing a value function using the Kantorovich-Rubinstein duality \cite{villani2008optimal}.
In this formulation, $W$ is approximated given a set of $k$-Lipschitz functions $f$ modeled as neural networks. To
ensure $f$ is $k$-Lipschitz, the weights of the discriminator are clipped to some range $[-c, c]$. In our work, we
adopt the Wasserstein GAN with gradient penalty (WGAN-GP) \cite{gulrajani2017improved}, which instead of clipping
network weights like in \cite{arjovsky2017wasserstein}, ensures the Lipschitz constraint by enforcing a soft
constraint on the gradient norm of the discriminator's output with respect to its input. Following
\cite{gulrajani2017improved}, our new objective then becomes

>>>>

\subsection{Adversarial Networks}
\label{sec:gans}
\nostarnote{check this paragraph as I added some more text}In machine learning literature, Generative Adversarial Networks 
(GANs)~\cite{goodfellow2014generative} represent a class of generative models based on game theory in which a generator network 
competes against an adversary. From a classification perspective, the generator network produces instances which actively attempt 
to `fool' the discriminator network. The goal is for the discriminator network to be able to distinguish between `true' 
instances coming from the dataset and `false' instances produced by the generator network.
%, and thus improve its classification accuracy.
In our case, conditioned on an image $I^D$, the generator is trained to produce an image to try and fool the 
discriminator, which is trained to distinguish between distorted and non-distorted underwater images. In the original
GAN formulation, our goal is to solve the minimax problem:

\begin{figure}
\vspace{2mm}
\centering
\begin{tabular}{p{1.7cm} p{1.7cm} p{1.7cm} p{1.7cm}}

@@

\end{tabular}
\caption{Paired samples of ground truth and distorted images generated by CycleGAN. Top row: Ground truth.
Bottom row: Generated samples.}
\label{fig:cgan_samples}
\vspace{-5mm}
\end{figure}

@@

Note for simplicity in notation, we will further omit $I^C \sim p_{train}$ and $I^D \sim p_{gen}$. In this formulation, 
the discriminator is hypothesized as a classifier with a sigmoid cross-entropy loss function, which in practice may lead
to issues  such as the vanish gradient and mode collapse. As shown by \cite{arjovsky2017towards}, as
the discriminator improves, the gradient of the generator vanishes, making it difficult or impossible to train. Mode
collapse occurs when the generator ``collapses'' onto a single point, fooling the discriminator with only one instance.
To illustrate the effect of mode collapse, imagine a GAN is being used to generate digits from the MNIST~\cite{lecun2010mnist} 
dataset, but it only generated the same digit. In reality, the desired outcome would be to generate a diverse collection of all 
the digits.
There have been a number of recent methods which hypothesize a different loss function for the discriminator
\cite{mao2016least,arjovsky2017wasserstein,gulrajani2017improved,zhao2016energy}. We focus on the Wasserstein GAN
(WGAN) \cite{arjovsky2017wasserstein} formulation, which proposes to use the Earth-Mover or \textit{Wasserstein-1}
distance $W$ by constructing a value function using the Kantorovich-Rubinstein duality \cite{villani2008optimal}.
In this formulation, $W$ is approximated given a set of $k$-Lipschitz functions $f$ modeled as neural networks. To
ensure $f$ is $k$-Lipschitz, the weights of the discriminator are clipped to some range $[-c, c]$. In our work, we
adopt the Wasserstein GAN with gradient penalty (WGAN-GP) \cite{gulrajani2017improved}, which instead of clipping
network weights like in \cite{arjovsky2017wasserstein}, ensures the Lipschitz constraint by enforcing a soft
constraint on the gradient norm of the discriminator's output with respect to its input. Following
\cite{gulrajani2017improved}, our new objective then becomes
